Doppler Imaging Fun

### 2019-06-12 tex/ms.tex
doing stuff in log space

Let $I(\lambda, x, y)$ be the Doppler-shifted intensity observed at wavelength $\lambda$ at
position $x, y$ on the surface of the star. We may express it as
%
\begin{align}
    \label{eq:IntensityUnif}
    I(\lambda, x, y) &= I_0\Big(\mathcal{D}^{-1}_{\beta(x, y)}(\lambda)\Big)
\end{align}
%
where $I_0$ is the (spatially uniform) spectrum in the original, unshifted frame
and 
%
\begin{align}
    \mathcal{D}^{-1}_{\beta(x, y)}(\lambda) &= \frac{\lambda}{1 + \beta(x, y)} \nonumber \\
                                            &\equiv \lambda_0
\end{align}
%
is the non-relativistic \emph{inverse} Doppler operator, which takes as input the shifted (observed)
wavelength $\lambda$ and returns the unshifted wavelength $\lambda_0$.
The quantity
%
\begin{align}
    \beta(x, y) &= \frac{v(x, y)}{c}
\end{align}
%
is the ratio of the radial velocity $v(x, y)$ at a point on the
surface to the speed of light.

If we Taylor expand Equation~(\ref{eq:IntensityUnif}) about $\beta = 0$, we obtain
%
\begin{align}
    \label{eq:TaylorUnifExplicit}
    I(\lambda, x, y) 
        &=
        I_0(\lambda_0) \Bigg|_{\beta=0}
        + 
        \frac{\mathrm{d}I_0(\lambda_0)}{\mathrm{d}\beta} \Bigg|_{\beta=0} \Delta\beta(x, y)
        + 
        \frac{\mathrm{d}^2I_0(\lambda_0)}{\mathrm{d}\beta^2} \Bigg|_{\beta=0} \Delta\beta(x, y)^2
        +
        ... 
\end{align}
%
The derivatives of the spectrum $I_0(\lambda_0)$ with respect to
$\beta$ are found by application of Fa\`a di Bruno's formula for taking high
order derivatives of the chain rule:
%
\begin{proof}{faa_di_bruno}
    \label{eq:dIdbeta}
    \frac{\mathrm{d}^n I_0(\lambda_0)}{\mathrm{d}\beta^n} \Bigg|_{\beta=0} &=
    \sum_{k=1}^n \frac{\mathrm{d}^k I(\lambda_0)}{\mathrm{d}\lambda_0^k} \lambda_0^k P_{nk}
\end{proof}
%
where
%
\begin{proof}{faa_di_bruno}
    \label{eq:Pnk}
    P_{nk} \equiv B_{n, k}\Bigg( \Big\{(-1)^j j! \Big\}_{j=1}^{n - k + 1} \Bigg)
\end{proof}
%
is an integer computed from the incomplete Bell polynomial $B_{n, k}$.
Given this result, and noting that $\lambda_0 = \lambda$ when $\beta = 0$,
we may re-write Equation~(\ref{eq:TaylorUnifExplicit}) as
%
\begin{align}
    \label{eq:TaylorUnifSum}
    I(\lambda, x, y) 
        &=
        I_0(\lambda)
        + 
        \sum_{n=1}^\infty
            \Bigg(
                \sum_{k=1}^n \frac{\mathrm{d}^k I(\lambda)}{\mathrm{d}\lambda^k} \lambda^k P_{nk} 
            \Bigg)
            \Bigg(
                \Delta\beta(x, y)^n
            \Bigg)
\end{align}
%
Assuming we only have access to spatially unresolved spectra of the star, the quantity
we observe is the integral of the intensity over the surface of the star, which we
denote $S$:
%
\begin{align}
    \label{eq:TaylorUnifSumIntegral}
    S(\lambda) 
        &=
        \pi I_0(\lambda)
        + 
        \sum_{n=1}^\infty
            \Bigg(
                \sum_{k=1}^n \frac{\mathrm{d}^k I(\lambda)}{\mathrm{d}\lambda^k} \lambda^k P_{nk} 
            \Bigg)
            \Bigg(
                \int{\Delta\beta(x, y)^n}\mathrm{d}\Omega
            \Bigg)
\end{align}
%
where $\int\mathrm{d}\Omega$ denotes the surface integral over the visible disk
of the star.

Note, importantly, that the terms in the first set of parentheses depend \emph{only} on the wavelength,
while the terms in the second set of parentheses depend \emph{only} on the position.
This Taylor expansion has therefore allowed us to deconvolve the effects of the Doppler
shift from the actual spectrum of the star.

>>>>

Let $I(\xi, x, y)$ be the Doppler-shifted intensity observed at log wavelength $\xi \equiv \ln\lambda$ at
sky-projected Cartesian position $x, y$ on the surface of the star. We may express it as
%
\begin{align}
    \label{eq:IntensityUnif}
    I(\xi, x, y) &= I_0(\xi_0)
\end{align}
%
where $I_0$ is the (spatially uniform) spectrum in the original, unshifted frame
and 
%
\begin{align}
    \xi_0 &= \xi - \alpha(x, y)
\end{align}
%
is the log wavelength in the unshifted frame. The quantity $\alpha$ is
computed from the formula for the relativistic Doppler shift and is equal to
%
\begin{align}
    \alpha(x, y) &\equiv \frac{1}{2}\ln\left( \frac{1 - \nicefrac{v}{c}}{1 + \nicefrac{v}{c}} \right)
\end{align}
%
where $v = v(x, y)$ is the radial velocity at a point on 
the surface of the star and $c$ is the speed of light.

If we Taylor expand Equation~(\ref{eq:IntensityUnif}) about $\alpha = 0$, we obtain
%
\begin{align}
    \label{eq:TaylorUnifExplicit}
    I(\xi, x, y) 
        &=
        I_0(\xi_0) \Bigg|_{\alpha=0}
        + 
        \frac{\mathrm{d}I_0(\xi_0)}{\mathrm{d}\alpha} \Bigg|_{\alpha=0} \Delta\alpha(x, y)
        + 
        \frac{1}{2}\frac{\mathrm{d}^2I_0(\xi_0)}{\mathrm{d}\alpha^2} \Bigg|_{\alpha=0} \Delta\alpha(x, y)^2
        +
        ... 
\end{align}
%
The derivatives of the spectrum $I_0(\xi_0)$ with respect to
$\alpha$ are computed by repeated application of the chain rule:
%
\begin{align}
    \frac{\mathrm{d}^nI_0(\xi_0)}{\mathrm{d}\alpha^n} &=
    (-1)^n\dfrac{\mathrm{d}^nI_0(\xi_0)}{\mathrm{d}\xi_0^n}
\end{align}
%
Given this result, and noting that $\xi_0 = \xi$ when $\alpha = 0$,
we may re-write Equation~(\ref{eq:TaylorUnifExplicit}) as
%
\begin{align}
    \label{eq:TaylorUnifSum}
    I(\xi, x, y) 
        &=
        I_0(\xi)
        +
        \sum_{n=1}^\infty
            \frac{(-1)^n}{n!}
            \frac{\mathrm{d}^nI_0(\xi)}{\mathrm{d}\xi^n}
            \Delta\alpha(x, y)^n
\end{align}
